### 2022-01-26 paper/Chapters/Evaluation.tex
First draft of outline
This also creates the file structure for each of the chapters, which is nice

\chapter{Evaluation} % Main chapter title

\label{Chapter6} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Testing Methodology}

\todosection

\section{Responsivity and Latency}

\todosection

\section{Performance}

\todosection

\section{Quality}

\todosection

\section{Summary}

\todosection