### 2022-01-31 paper/Chapters/Background.tex

\chapter{Background} % Main chapter title

\label{Chapter2} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Specialization of Computers}

\todosection

\subsection{Power of the Desktop}

\todosection

\subsection{Convenience of the Laptop}

\todosection

\subsection{Rise of the Gaming Laptop}

\todosection

\section{Thin Clients}

\todosection

\section{Application to Modern Day}

\todosection