### 2022-01-31 paper/Chapters/Introduction.tex

\chapter{Introduction} % Main chapter title

\label{Chapter1} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Power and Portability}

\todosection

\section{Purpose of this Thesis}

\todosection

\section{Thesis Overview}

\todosection