### 2022-01-31 paper/Chapters/Software.tex

\chapter{Developing the Software} % Main chapter title

\label{Chapter5} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Requirements}

\todosection

\section{Potential Avenues}

\todosection

\subsection{NVIDIA GameStream}

\todosection

\subsection{Moonlight}

\todosection

\section{Developing for ARM}

\todosection