### 2022-01-31 paper/Chapters/StateOfTheArt.tex

\chapter{State of the Art} % Main chapter title

\label{Chapter3} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Introduction}

\todosection

\section{Hardware Solutions}

\todosection

\subsection{Thin Clients}

\todosection

\subsection{Web Clients}

\todosection

\section{Software Solutions}

\todosection

\subsection{Remote Desktop Protocol}

\todosection

\subsection{Virtual Network Computing}

\todosection

\subsection{Chrome Remote Desktop}

\todosection

\subsection{Secure Shell Protocol}

\todosection