### 2022-03-22 paper/Appendices/AppendixB.tex

\chapter{Bill of Materials} % Main appendix title

\label{AppendixB} % For referencing this appendix elsewhere, use \ref{AppendixA}

\begin{longtable}{|>{\raggedright\arraybackslash}p{0.2\linewidth}|c|l|l|}%
  \hline
  \label{tab:BillOfMaterials}
  \bfseries Reference \newline Designator& \bfseries Count & \bfseries Manufacturer's Number &\bfseries Value % Columns
  \csvreader[respect percent=true, respect underscore=true, respect sharp=true, table head=\toprule, table foot=\bottomrule]{Figures/kicad/lathrum_thesis_bom.csv}{}% Data
  {\\\hline\csvcoli&\csvcolii&\csvcolv&\csvcoliii}% Rows
  \\\hline
\end{longtable}