### 2022-04-01 paper/Chapters/Software.tex
Start developing for ARM section

\chapter{Developing the Software} % Main chapter title

\label{Chapter5} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\todosection

\section{Requirements}\label{sec:SoftwareRequirements}

Much like the hardware requirements in section \ref{sec:HardwareRequirements}, the primary requirement of the software is to be performant enough to provide the user with a positive experience.
This is broken down to include being able to stream and decode video over the internet with as little latency as possible, stream inputs from the user to the remote computer, and retain visual fidelity while the host machine is under heavy load.
While much of the latency between the host and client is due to the network both devices are connected to, there is still some optimizations that can be made to reduce the time between data being send and images being displayed on the screen.
Similarly, while the rate at which data is transferred between the host and client cannot exceed the speed of the slowest internet connection of the two devices, there can still be a balance between the bandwidth-saving but lower quality streaming seen in Chrome Remote Desktop (\ref{subsec:ChromeRemoteDesktop}), and the local network but higher quality Nvidia Shield (\ref{subsec:NvidiaShieldAndGameStream}).
To accomplish this, a compression method similar to CRD will be used, but it will not be bound by the traditional web browser which limit CRD due to the software solution being developed as a native application.

As a secondary objective, it would also be beneficial for the software to be able to run on a variety of different hardware platforms in the event the hardware is determined to be infeasible to develop, as discussed in Section \ref{subsec:HardwareCost}.

\section{Potential Avenues}\label{sec:PotentialAvenues}

As with any software project, there are many potential ways to solve the problem.
In many cases, the easiest place to start is to begin by ruling out potential avenues that are not feasible.
For example, as evident by the number of existing solutions that attempt to solve this problem (Chapter \ref{Chapter3}), taking into special account Nvidia's recent foray into remote gaming with the Shield (\ref{subsec:NvidiaShieldAndGameStream}), it is not reasonable to attempt to create a brand new software solution from scratch with the time and resources available.
It has taken Nvidia, a company worth hundreds of billions of dollars, years to get to the point they are at now \cite{brown_2013}.
Developing a rival protocol or new software platform is simply not possible for this project.
Instead, given that Nvidia's GameStream technology is the current leader in the realm of high-performance remote computing, the best way to build a software product that meets the requirements laid out in Section \ref{sec:ResearchQuestions} is to build upon what already exists.

\subsection{NVIDIA GameStream}\label{subsec:NVIDIAGameStream}

\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,inline]{Should this subsection and the following subsection be top level sections? The organization of the headers is a little strange, but the content flows pretty well}

Unfortunately, the Nvidia GameStream protocol is both proprietary and a closed source project.
Very little is officially known about the protocol, other than it uses specific video encoders built into Nvidia Graphics cards from the 600 series from 2012 and newer \cite{gamestream_userguide}.
Apart from that, everything has been researched and reverse engineered by the community.
A GameStream session is started by first connecting to a host computer which has enabled the GameStream service through Nvidia's GeForce Experience application.
Once the service is enabled, the computer will listen on port \emph{47989}\todoquestion{Should this be italicized? quoted? plaintext?} for the following web requests:
\begin{itemize}
  \item /serverinfo
  \item /pair
  \item /unpair
  \item /applist
  \item /launch
  \item /resume
  \item /cancel
\end{itemize}
Once a session has been started with a \emph{/launch} request, raw data covering video, audio, and input is passed back and forth over a number of ports until the session is terminated.
An end-to-end example of the protocol's usage is as follows:

\begin{enumerate}
  \item A user knows that a capable host computer is ready to stream at the address \emph{192.168.1.123}.
        \begin{enumerate}
          \item This can be confirmed by sending a HTTP GET request \emph{/serverinfo} to the host computer on port \emph{47989}. For example: \newline\emph{http://192.168.1.123:47989/serverinfo?uniqueid=1234}
        \end{enumerate}
  \item The user can make a request to \emph{/pair} with a number of parameters to authorize itself with the host.
        \begin{enumerate}
          \item Pairing is a two step-process.
                First, a request to pair is made, at which point the host machine will display a four digit code on it's screen.
                Then the client must send that code back to the host as a second form of authentication.
          \item The authentication created by this pairing process lasts until either the client makes an \emph{/unpair} request with it's unique ID or a user on the host machine unpairs the client.
        \end{enumerate}
  \item The user can make a request to \emph{/applist} to get a list of applications that the host is able to stream.
  \item The user makes a request to \emph{/launch} with a number of parameters to instruct the host to create a session with the given settings and launch the specified application.
        \begin{enumerate}
          \item The configuration options given include the application to launch, the video resolution to stream, audio setup, input mapping, and multiple forms of authentication.
        \end{enumerate}
  \item If the user unexpectedly or accidentally leaves a session without closing it, they can make a request to \emph{/resume} to instruct the host to reopen and resume the session the previous session.
  \item Once the user is finished, they make a request to \emph{/cancel} to instruct the host to close the session.
  \item If the user is completely finished with the host, they can return it to the state they found it in by making a request to \emph{/unpair} with their unique ID, unauthorizing themselves with the host.
\end{enumerate}

The most difficult part of this process is handling the transfer of data after a session has begun.
Once a session has launched, data is sent directly between the host and client over a number of ports without going through an easy to understand web protocol.
This is where the community project Moonlight steps in.

\subsection{Moonlight}\label{subsec:Moonlight}

Moonlight is an unofficial third-party open-source client for the Nvidia GameStream protocol \cite{moonlight}.
The project is structured as a core library written in \emph{C} with a number of community built clients built on top of it for various platforms.
In essence, it turns the process described in the previous section (\ref{subsec:NVIDIAGameStream}) into a usable programming interface.
By reverse engineering the protocol originally implemented by the Nvidia Shield, Moonlight enables the development of a GameStream client for almost any operating system.
The Moonlight community has already developed fantastic clients for many platforms such as Windows, Mac, Linux, Android, iOS, Web, and even other gaming consoles.
All that is needed for the purposes of this project is to make tweaks to best suit the CM4 that drives the hardware and develop a suite of tools to test it's functionality.

\section{Developing for ARM}\label{sec:DevelopingForARM}

The CM4 runs on a 64-bit ARM Cortex-A72 processor \cite{rpi_cm4}, meaning that applications must be compiled for the \emph{ARM} architecture in order for the CPU to be able to run it.
If a developer were to write a program on their \emph{x86} architecture Windows desktop computer, they could compile and run their program on any other Windows computer running on the same \emph{x86} architecture.
They could also compile their application for other operating systems, but the target machine must also be running the same architecture.
A computer running on another architecture such as \emph{ARM} is not able to run a program compiled for \emph{x86} machines, and a computer running on \emph{x86} is not able to run a program compiled for \emph{ARM} machines or any other architecture.

In order to run the code for the project, given that the CM4 is the only \emph{ARM} computer available for this project, the code must either be compiled on the CM4 or built in a cross-compilation environment on the \emph{x86} machine.
Using a cross-compilation environment would be overall quicker than that developing on the CM4 itself since it can use the full power of a desktop PC.
Thankfully, setting up a cross-compilation environment for projects written in \emph{C} is straightforward.
\emph{CMAKE}, a common compilation tool for \emph{C}, has built-in support for compiling for different target architectures with easy to use libraries.
What isn't as straightforward however, is building a toolchain to compile the rest of the code for \emph{ARM}.

In order to create the most straightforward experience possible for someone using the hardware for this project, a custom operating system image is created that has all the dependencies, libraries, and applications pre-installed.
Manjaro provides good tools for building and managing custom images of the Manjaro operating system for \emph{ARM} architectures \cite{manjaro_arm_tools}.
The only drawback is that it must be run an \emph{ARM} processor, and is much more difficult to run from another architecture.
To make this process easier to develop, a toolchain was developed that could be run on the CM4 to automatically compile all the necessary code using \emph{cmake}, package all the necessary files using \emph{makepkg}, and build a fresh OS image that is ready to flash.
This three step script allows the developer to pull the git repository of the project, run a set up script, and then run the build script to compile everything into an immediately deployable OS image.
The following excerpt details the last step of the process, building the OS image that includes the custom application.

\begin{lstlisting}[style=custombash,firstnumber=72,title=Excerpt from \emph{/code/build.sh}]
pkgname=$(sed -n 's/^pkgname=//p' ./pkgbuild/PKGBUILD)
pkgver=$(sed -n 's/^pkgver=//p' ./pkgbuild/PKGBUILD)
pkgrel=$(sed -n 's/^pkgrel=//p' ./pkgbuild/PKGBUILD)
# Copy local package to cache to be installed into new image
sudo cp ./pkgbuild/$pkgname-$pkgver-$pkgrel-aarch64.pkg.tar.zst /var/cache/pacman/pkg/
# Build image
sudo buildarmimg -d rpi4 -e thesis -i $pkgname-$pkgver-$pkgrel-aarch64.pkg.tar.zst
read -s -n 1 -p "Image built, press a key to continue (for sudo)"
# Remove local package from cache
sudo rm /var/cache/pacman/pkg/$pkgname-$pkgver-$pkgrel-aarch64.pkg.tar.zst -f
# Move build image to build folder
sudo mv /var/cache/manjaro-arm-tools/img/* ./build/
\end{lstlisting}

This script pulls the package's name, version, and release number from the \emph{PKGBUILD} file generated from the packaging step and installs it into the newly created image, which is then moved into the project's build directory.
This image file is then ready to be flashed and installed like any other operating system.

\section{Developing Testing and Measurement Tools}\label{sec:DevelopingTestingAndMeasurementTools}

In order to evaluate the performance of the project, a suite of tools must be developed to test both the newly developed application and the existing solutions.
To do this, two sets of data will be collected: First, data directly from the application that logs statistics such as the number of frames rendered, dropped, or rendered, and data that runs separately from the application that logs the amount of time it takes for data that is from the client to be realized as a change on the screen.
The first set of data is important to track the raw performance of the application, statistics like Frames Per Second (FPS)

\begin{lstlisting}[style=customc,firstnumber=23,title=Excerpt from \emph{/src/video/stats.h}]
typedef struct _VIDEO_STATS {
  uint32_t receivedFrames;
  uint32_t decodedFrames;
  uint32_t renderedFrames;
  uint32_t totalFrames;
  uint32_t networkDroppedFrames;
  uint32_t totalDecodeTime;
  uint32_t totalRenderTime;
  uint32_t lastRtt; // Rtt = Round Trip Time
  uint32_t lastRttVariance;
  float receivedFps;
  float decodedFps;
  float renderedFps;
  uint32_t measurementStartTimestamp;
} VIDEO_STATS;
\end{lstlisting}