### 2022-04-05 paper/Chapters/Introduction.tex

\chapter{Introduction} % Main chapter title

\label{Chapter1} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Power and Portability}

As computers continue to grow to become more and more integrated into our daily lives, purchasing a computer has always been a balance.
Whether its deciding between something cheap or something long-lasting, a machine that is pre-built or a machine that is custom tailored and custom built, or, a tradeoff that has become much more common in recent years, something powerful or something portable.
As with anything, compromises can always be made and a middle ground may be found, but the only way to get the best of both worlds currently is to pay a price premium.
If someone were to look for a computer that is powerful, long-lasting, and upgradable, they would be directed towards a desktop computer.
However, if they needed something with the same power but portable, their only option would be to purchase a laptop costing much more than its desktop counterpart and often with limitations in upgradeability down the line.
Without purchasing both an expensive desktop and an expensive laptop, there currently is no way to get the benefits of both power and portability.

\section{Purpose of this Thesis}

The primary goal of this thesis is to create a portable hardware and software solution that can be used to control a desktop computer remotely while leveraging all of it's power.
There are existing software solutions that give a user remote control of their desktop computer from a different location, but many of them fail to provide a high-quality and performant solution to working with highly demanding programs that put the desktop computer under intense load.
Once attempting to run an intensive application such as rendering video, training Artificial Intelligence, or developing a video game, existing solutions fail to keep up.

In other words, this thesis seeks to provide a user who already owns a powerful desktop computer with a mobile solution that gives them the ability to leverage it's power remotely.

\section{Thesis Overview}

This thesis is split into seven chapters.
Chapter \ref{Chapter2} introduces background knowledge surrounding existing computers, the shift to a more mobile computing world, and the modern day applications for this technology.
Chapter \ref{Chapter3} discusses the current the state of the art in terms of existing solutions for remote computer access, and presents a list of questions that this thesis seeks to answer.
Chapters \ref{Chapter4} and \ref{Chapter5} detail the requirements, design, and production of the hardware and software portions of the project respectively.
This design is tested against the requirements and the existing solutions in Chapter \ref{Chapter6}.
Finally, Chapter \ref{Chapter7} concludes the thesis with a summary of the project's innovations, limitations, and future research.